An Embedding Framework for Consistent Polyhedral Surrogates

### 2022-01-11 poly-embed-journal/poly-embed-journal.tex
minor edits in section 4

Theorem \ref{thm:calibrated-separated} will show that calibration is equivalent to a geometric feature, which we call \emph{separation}, of a link function $\psi$.
Recall that for indirect elicitation, any point $u \in \Gamma(p)$ must link to a report $\psi(u) \in \gamma(p)$.
(In terms of losses, $u$ minimizing expected $L$-loss implies that $\psi(u)$ minimizes expected $\ell$-loss, with respect to $p$.)
The idea of separation is that points in the neighborhood of $u$ must also link to to a report in $\gamma(p)$.
Furthermore, there must be a lower bound $\epsilon$ on the size of any such neighborhood.

@@

%\raf{Equivalence blah blah blah... proof appears in Appendix~\ref{sec:equiv-sep-calib}.}
\begin{restatable}{theorem}{calibratedseparated} \label{thm:calibrated-separated}
  Let polyhedral surrogate $L:\reals^d \to \reals^\Y_+$, discrete loss $\ell:\R\to\reals^\Y_+$, and link $\psi:\reals^d\to\R$ be given.
  Then $(L,\psi)$ is calibrated with respect to $\ell$ if and only if
  $\psi$ is $\epsilon$-separated with respect to $L$ and $\ell$ for some
  $\epsilon>0$.
\end{restatable}

\bo{Say anything about proof of above theorem?}
To prove Theorem \ref{thm:link-main}, it now suffices to show that for any polyhedral $L$ embedding some $\ell$, there exists a \emph{separated} link $\psi$ with respect to $L$ and $\ell$.
This is given by Construction ~\ref{const:eps-thick-link} below.

@@

For \emph{(a)}, begin by linking each embedding point back to its original report.
Now we must determine $\psi(u)$ for non-embedding points.
%Symbolically, if $u^* \in \argmin_u \inprod{p}{L(u)} = \Gamma(p)$, then $\psi(u^*) \in \argmin_r \inprod{p}{\ell(r)} = \gamma(p)$.
The challenge is that we may have $u \in \Gamma(p) \cap \Gamma(p')$.
Because $u$ minimizes expected surrogate loss for both $p$ and $p'$, the link must satisfy $\psi(u) \in \gamma(p) \cap \gamma(p')$.
It is not even clear \emph{a priori} that these sets intersect.
We use the definition of embedding and elicitation results, discussed in \S~\ref{sec:min-rep-sets}, to show that for each such $u$ there exists $r \in \R$ such that $\Gamma_u \subseteq \gamma_r$, i.e. any $p$ satisfying $u \in \Gamma(p)$ also satisfies $r \in \gamma(p)$.
This implies that if $u \in \Gamma(p) \cap \Gamma(p')$, then $r \in \gamma(p) \cap \gamma(p')$, so we may safely choose $\psi(u) = r$.

For \emph{(b)}, we show that this link can be ``thickened'' by some positive $\epsilon$, as described next.
Consider an optimal report set, i.e. set of the form $U = \Gamma(p) = \argmin_u \inprod{p}{L(u)}$.
By indirect elicitation, $\psi$ is already correct on $U$.
Now, we ``thicken'' $U$ to obtain $U_{\epsilon} = \{u : \|u - U\| \leq \epsilon\}$.
Then we require that all points in $U_{\epsilon}$ are linked to some element of $\gamma(p) = \argmin_r \inprod{p}{\ell(r)}$.
For $\epsilon > 0$, this directly implies separation.

@@

Given $u \in U \cap U' \cap \dots$, we use $\Psi(u)$ to denote the remaining legal choices for $\psi(u)$ after imposing the requirements for each such set $U,U'$, etc.
The key claim is that, for small enough $\epsilon > 0$, $\Psi(u)$ is nonempty: there exists at least one legal report to link to.
This claim follows from a geometric result (Lemma \ref{lemma:thick-empty}) that, for all small enough $\epsilon$, a subset of thickenings $U_{\epsilon}$ intersect if and only if the $U$ sets themselves intersect.
When they do intersect, indirect elicitation implies that there exists a legal choice of link for the intersection of the thickenings.
It is also important that, by Lemma~\ref{lem:polyhedral-range-gamma}, for polyhedral surrogates there are only finitely many sets of the form $U = \Gamma(p)$.
This yields a single uniform smallest $\epsilon$ such that the key claim is true for all $u \in \reals^d$.
%Given a well-defined thickened link $\psi$, calibration is almost immediate because, given $p$, any report $u$ linking to a suboptimal $r = \psi(u)$ satisfies $\inf_{u^* \in \Gamma(p)} \|u^* - u\| \geq \epsilon$.

@@

Applying the $\epsilon$-thickened link construction additionally enables one to verify the consistency of a proposed link.
For a given $\epsilon$ and norm $\|\cdot\|$, suppose one follows the routine of Construction~\ref{const:eps-thick-link} until the last step in which values for the link are selected.
Once possible choices for this $\epsilon$-thickened set-valued function are available, we can see if the proposed link aligns with these possible choices: if so, then the proposed link is calibrated.

\bo{Should we comment on computing the $\epsilon$?}

\botodo{Commented out Proposition: converse, that every calibrated link from a polyhedral surrogate can be produced by our construction.}
%\paragraph{A converse to Theorem \ref{thm:thickened-separated}.}
%Perhaps surprisingly, one can also show that \emph{every} calibrated link from a polyhedral surrogate to a discrete loss is produced by Construction \ref{const:eps-thick-link}, for some tiebreaking rule.
%The proof appears in \bo{TODO}.
%
%\begin{proposition}
%  Let polyhedral surrogate $L:\reals^d \to \reals^\Y_+$ and discrete loss $\ell:\R\to\reals^\Y_+$ be given.
%  Then for every calibrated link $\psi$, there exists $\epsilon>0$ such that Construction~\ref{const:eps-thick-link} yields $\psi$.
%\end{proposition}

>>>>

Theorem \ref{thm:calibrated-separated} shows that calibration is equivalent to a geometric condition, which we call \emph{separation}, of a link function $\psi$.
Recall that for indirect elicitation, any point $u \in \Gamma(p)$ must link to a report $\psi(u) \in \gamma(p)$.
(In terms of losses, $u$ minimizing expected $L$-loss implies that $\psi(u)$ minimizes expected $\ell$-loss, with respect to $p$.)
The idea of separation is that points in the neighborhood of $u$ must also link to to a report in $\gamma(p)$.
Furthermore, there must be a uniform lower bound $\epsilon$ on the size of any such neighborhood.

@@

\begin{restatable}{theorem}{calibratedseparated} \label{thm:calibrated-separated}
  Let polyhedral surrogate $L:\reals^d \to \reals^\Y_+$, discrete loss $\ell:\R\to\reals^\Y_+$, and link $\psi:\reals^d\to\R$ be given.
  Then $(L,\psi)$ is calibrated with respect to $\ell$ if and only if
  $\psi$ is $\epsilon$-separated with respect to $L$ and $\ell$ for some
  $\epsilon>0$.
\end{restatable}
\bo{Say anything about proof of above theorem?}
\raf{We definitely should for the journal submission!  Punt for now.}

To prove Theorem \ref{thm:link-main}, it now suffices to show that for any polyhedral $L$ embedding some $\ell$, there exists a \emph{separated} link $\psi$ with respect to $L$ and $\ell$.
This is given by Construction ~\ref{const:eps-thick-link} below.

@@

For \emph{(a)}, begin by linking each embedding point back to its original report.
Now we must determine $\psi(u)$ for non-embedding points.
%Symbolically, if $u^* \in \argmin_u \inprod{p}{L(u)} = \Gamma(p)$, then $\psi(u^*) \in \argmin_r \inprod{p}{\ell(r)} = \gamma(p)$.
The challenge is that we may have $u \in \Gamma(p) \cap \Gamma(p')$.
Because $u$ minimizes expected surrogate loss for both $p$ and $p'$, the link must satisfy $\psi(u) \in \gamma(p) \cap \gamma(p')$.
It is not even clear \emph{a priori} that these sets intersect.
We use the definition of embedding and elicitation results, discussed in \S~\ref{sec:min-rep-sets}, to show that for each such $u$ there exists $r \in \R$ such that $\Gamma_u \subseteq \gamma_r$, i.e. any $p$ satisfying $u \in \Gamma(p)$ also satisfies $r \in \gamma(p)$.
This implies that if $u \in \Gamma(p) \cap \Gamma(p')$, then there exists $r \in \gamma(p) \cap \gamma(p')$, so we may safely choose $\psi(u) = r$.

For \emph{(b)}, we show that this link can be ``thickened'' by some positive $\epsilon$, as described next.
Consider an optimal surrogate report set, i.e. set of the form $U = \Gamma(p) = \argmin_u \inprod{p}{L(u)}$.
By indirect elicitation, $\psi$ is already correct on $U$.
Now, we ``thicken'' $U$ to obtain $U_{\epsilon} = \{u : \|u - U\| \leq \epsilon\}$.
Then we require that all points in $U_{\epsilon}$ are linked to some element of $\gamma(p) = \argmin_r \inprod{p}{\ell(r)}$.
For $\epsilon > 0$, this directly implies separation.

@@

Given $u \in U \cap U' \cap \dots$, we use $\Psi(u)$ to denote the remaining legal choices for $\psi(u)$ after imposing the requirements for each such set $U,U'$, etc.
The key claim is that, for small enough $\epsilon > 0$, $\Psi(u)$ is nonempty: at least one legal value for $\psi(u)$ remains.
This claim follows from a geometric result (Lemma \ref{lemma:thick-empty}) that, for all small enough $\epsilon$, a subset of thickenings $U_{\epsilon}$ intersect if and only if the $U$ sets themselves intersect.
When they do intersect, indirect elicitation implies that there exists a legal choice of link for the intersection of the thickenings.
It is also important that, by Lemma~\ref{lem:polyhedral-range-gamma}, for polyhedral surrogates there are only finitely many sets of the form $U = \Gamma(p)$.
This yields a single uniform smallest $\epsilon$ such that the key claim is true for all $u \in \reals^d$.
%Given a well-defined thickened link $\psi$, calibration is almost immediate because, given $p$, any report $u$ linking to a suboptimal $r = \psi(u)$ satisfies $\inf_{u^* \in \Gamma(p)} \|u^* - u\| \geq \epsilon$.

@@

Applying the $\epsilon$-thickened link construction additionally enables one to verify the consistency of a proposed link $\psi^*$.
For a given $\epsilon$ and norm $\|\cdot\|$, suppose one follows the routine of Construction~\ref{const:eps-thick-link} until the last step in which values for the link $\psi$ are selected.
Instead, we can simply test whether the proposed link values are contained in the valid choices, i.e., if $\psi^*(u) \in \Psi(u)$ for all $u\in\reals^d$.
If so, then the proposed link $\psi^*$ is calibrated.

\bo{Should we comment on computing the $\epsilon$?}
\raf{Don't recall what we have to say... but this would be a great spot to do so.  Or at least make the point that the regret bounds (below) get better the higher $\epsilon$ is.}

\bo{Add for journal submission: converse, that every calibrated link from a polyhedral surrogate can be produced by our construction.}
%\paragraph{A converse to Theorem \ref{thm:thickened-separated}.}
%Perhaps surprisingly, one can also show that \emph{every} calibrated link from a polyhedral surrogate to a discrete loss is produced by Construction \ref{const:eps-thick-link}, for some tiebreaking rule.
%The proof appears in \bo{TODO}.
%
%\begin{proposition}
%  Let polyhedral surrogate $L:\reals^d \to \reals^\Y_+$ and discrete loss $\ell:\R\to\reals^\Y_+$ be given.
%  Then for every calibrated link $\psi$, there exists $\epsilon>0$ such that Construction~\ref{const:eps-thick-link} yields $\psi$.
%\end{proposition}
